---
title: "CDC SVI // Clean ACS Data"
author: "Enterprise Analytics"
date: "`r paste0('Last Updated: ', format(Sys.time(), '%d %B, %Y')) `"
keywords: 
  - ACS
  - CDC SVI
output: 
  eaverse::ea_report:
    analyst: 
    - "P. Walker"
    email: "philip.walker@conehealth.com"
    draft: true
    code_folding: "show" # one of "show", "hide", or "none"
    code_download: FALSE # TRUE to allow the Rmd to be downloaded or FALSE
# knit: eaverse::knit_ea_letter
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
#knitr::opts_chunk$set(fig.align = 'center')
library(tidyverse)
library(kableExtra)
library(eaverse)
library(patchwork)
# To Turn off Scientific Notation Use this.
options(scipen = 999)

# This cleans up numbers to add decimal places
knitr::knit_hooks$set(inline = clean_numbers,
                      dropdown = knit_dropdown,
                      chunk = ea_hooks)
```

```{r}

load(file = here::here("data", "ACS_codes.rda"))

```

# Group: Overall{.tabset .tabset-pills}

## Total Population
```{r, fig.width=12, fig.height=5, out.width="100%"}

load(here::here("data", "E_TOTPOP.rda"))
objectNM <- unique(longACS$SVI_metric)
assign(objectNM, longACS, pos=1)

# E_TOTPOP <-
#   E_TOTPOP %>% 
#   mutate(PCT = ntile(values, 100)/100)

# Perct <- E_TOTPOP %>% 
#    ggplot(aes(PCT))+
#    geom_histogram(bins=10, fill='#69b3a2', color='white') + 
#    scale_x_continuous(labels = scales::percent_format())+
#    labs(subtitle = "By Count at Percentile Group")

ValueP <- E_TOTPOP %>% 
  ggplot(aes(values))+
  geom_histogram(bins=10, fill='#69b3a2', color='white') + 
  # scale_x_continuous(labels = scales::percent_format())+
  labs(
       subtitle = "By Count of Counties in at value range")

# Perct + 
  ValueP +
  plot_annotation(title = "Distribution of Population")

```


## Housing Units Per Blockgroup

```{r, fig.width=12, fig.height=5, out.width="100%"}
# rm(longACS)
load(here::here("data", "E_HU.rda"))
objectNM <- unique(longACS$SVI_metric)
assign(objectNM, longACS, pos=1)

# E_HU <-
#   E_HU %>% 
#   mutate(PCT = ntile(values, 100)/100)

ValueP <- E_HU %>% 
  ggplot(aes(values))+
  geom_histogram(bins=10, fill='#69b3a2', color='white') + 
  # scale_x_continuous(labels = scales::percent_format())+
  labs(title = "Distribution of Housing Units",
       subtitle = "By Count of Counties in at value range")

ValueP

```

## Household Estimates Per Blockgroup

```{r, fig.width=12, fig.height=5, out.width="100%"}
# rm(longACS)
load(here::here("data", "E_HH.rda"))
objectNM <- unique(longACS$SVI_metric)
assign(objectNM, longACS, pos=1)

# E_HH <-
#   E_HH %>% 
#   mutate(PCT = ntile(values, 100)/100)

ValueP <- E_HH %>% 
  ggplot(aes(values))+
  geom_histogram(bins=10, fill='#69b3a2', color='white') + 
  # scale_x_continuous(labels = scales::percent_format())+
  labs(title = "Distribution of Household Estimates",
       subtitle = "By Count at value range")

ValueP

```



# Group: SES  {.tabset .tabset-pills}

## Persons below poverty estimate Per Blockgroup

```{r, fig.width=12, fig.height=5, out.width="100%"}
# rm(longACS)
load(here::here("data", "E_POV.rda"))
objectNM <- unique(longACS$SVI_metric)
assign(objectNM, longACS, pos=1)

E_POV <-
  E_POV %>% 
  select(-vars) %>% 
  group_by_at(vars(-values)) %>%
  summarize(SumPov = sum(values)) %>% 
  ungroup() %>%   
  left_join(., 
            E_TOTPOP %>% select(NAME, pop = values), 
            by = c("NAME")) %>% 
  mutate(PercPopBelow = SumPov/pop,
         PCT = ntile(PercPopBelow, 100)/100)

ValueP <- E_POV %>% 
  ggplot(aes(PercPopBelow))+
  geom_histogram(bins=10, fill='#69b3a2', color='white') + 
  scale_x_continuous(labels = scales::percent_format())+
  labs(subtitle = "Percent of Population Estimated Below Poverty",
       y=NULL, x=NULL)

ValueP +
  plot_annotation(title = "Distribution of Persons below poverty estimate")

```



## Unemployment

```{r, fig.width=12, fig.height=5, out.width="100%"}
# rm(longACS)
load(here::here("data", "E_UNEMP.rda"))
objectNM <- unique(longACS$SVI_metric)
assign(objectNM, longACS, pos=1)

E_UNEMP <-
  E_UNEMP %>% 
  left_join(., 
            E_TOTPOP %>% select(NAME, pop = values), 
            by = c("NAME")) %>% 
  mutate(PercPopUnEmp = values/pop,
         PCT = ntile(PercPopUnEmp, 100)/100)

ValueP <- E_UNEMP %>% 
  ggplot(aes(PercPopUnEmp))+
  geom_histogram(bins=10, fill='#69b3a2', color='white') + 
  scale_x_continuous(labels = scales::percent_format())+
  labs(subtitle = "Percent of Population Estimated Below Poverty",
       y=NULL, x=NULL)

ValueP +
  plot_annotation(title = "Distribution of Persons below poverty estimate")

```


## Income Level
Per Capita Income

```{r, fig.width=12, fig.height=5, out.width="100%"}
# rm(longACS)
load(here::here("data", "E_PCI.rda"))
objectNM <- unique(longACS$SVI_metric)
assign(objectNM, longACS, pos=1)

E_PCI <-
  E_PCI %>% 
  mutate(PCT = ntile(values, 100)/100) %>% 
  mutate(PCT = 1 - PCT)

ValueP <- E_PCI %>% 
  ggplot(aes(values))+
  geom_histogram(bins=10, fill='#69b3a2', color='white') + 
  # scale_x_continuous(labels = scales::percent_format())+
  labs(#subtitle = "Percent of Population Estimated Below Poverty",
       y=NULL, x=NULL)

ValueP +
  plot_annotation(title = "Distribution of Per Captia Income")


```



## No High School Diploma

```{r, fig.width=12, fig.height=5, out.width="100%"}
# rm(longACS)
load(here::here("data", "E_NOHSDP.rda"))
objectNM <- unique(longACS$SVI_metric)
assign(objectNM, longACS, pos=1)

E_NOHSDP <-
  E_NOHSDP %>% 
  select(-vars) %>% 
  group_by_at(vars(-values)) %>%
  summarize(SumNoDP = sum(values)) %>% 
  ungroup() %>% 
  left_join(., 
            E_TOTPOP %>% group_by(tract) %>% summarise(pop = sum(values)), 
            by = c("tract")) %>% 
  mutate(PercNoDP = SumNoDP/pop,
         PCT = ntile(PercNoDP, 100)/100)

ValueP <- E_NOHSDP %>% 
  ggplot(aes(PercNoDP))+
  geom_histogram(bins=10, fill='#69b3a2', color='white') + 
  scale_x_continuous(labels = scales::percent_format())+
  labs(#subtitle = "Percent of Population Estimated Below Poverty",
       y=NULL, x=NULL)

ValueP +
  plot_annotation(title = "Distribution of Percentage of Block Group Without HS Diploma")


```



# Group: Household Composition  {.tabset .tabset-pills}

## Older than 65

```{r, fig.width=12, fig.height=5, out.width="100%"}
# rm(longACS)
load(here::here("data", "E_AGE65.rda"))
objectNM <- unique(longACS$SVI_metric)
assign(objectNM, longACS, pos=1)

E_AGE65 <-
  E_AGE65 %>% 
  select(-vars) %>% 
  group_by_at(vars(-values)) %>%
  summarize(SumOlder = sum(values)) %>% 
  ungroup() %>% 
  left_join(., 
            E_TOTPOP %>% select(NAME, pop = values), 
            by = c("NAME")) %>% 
  mutate(PercPopOlder= SumOlder/pop,
         PCT = ntile(PercPopOlder, 100)/100)

ValueP <- E_AGE65 %>% 
  ggplot(aes(PercPopOlder))+
  geom_histogram(bins=10, fill='#69b3a2', color='white') + 
  scale_x_continuous(labels = scales::percent_format())+
  labs(#subtitle = "Percent of Population Estimated Below Poverty",
       y=NULL, x=NULL)

ValueP +
  plot_annotation(title = "Distribution Block Group Above 65")


```



## Younger than 17

```{r, fig.width=12, fig.height=5, out.width="100%"}
# rm(longACS)
load(here::here("data", "E_AGE17.rda"))
objectNM <- unique(longACS$SVI_metric)
assign(objectNM, longACS, pos=1)

E_AGE17 <-
  E_AGE17 %>% 
  select(-vars) %>% 
  group_by_at(vars(-values)) %>%
  summarize(SumYoung = sum(values)) %>% 
  ungroup() %>% 
  left_join(., 
            E_TOTPOP %>% select(NAME, pop = values), 
            by = c("NAME")) %>% 
  mutate(PercPopYounger= SumYoung/pop,
         PCT = ntile(PercPopYounger, 100)/100)

ValueP <- E_AGE17 %>% 
  ggplot(aes(PercPopYounger))+
  geom_histogram(bins=10, fill='#69b3a2', color='white') + 
  scale_x_continuous(labels = scales::percent_format())+
  labs(#subtitle = "Percent of Population Estimated Below Poverty",
       y=NULL, x=NULL)

ValueP +
  plot_annotation(title = "Distribution Block Group Below 17")


```

## Older than 5 with a disability

```{r, fig.width=12, fig.height=5, out.width="100%"}
# rm(longACS)
load(here::here("data", "E_DISABL.rda"))
objectNM <- unique(longACS$SVI_metric)
assign(objectNM, longACS, pos=1)

E_DISABL <-
  E_DISABL %>% 
  select(-vars) %>% 
  group_by_at(vars(-values)) %>%
  summarize(SumDisab = sum(values)) %>% 
  ungroup() %>% 
  left_join(., 
            E_TOTPOP %>% group_by(tract) %>% summarise(pop = sum(values)), 
            by = c("tract")) %>% 
  mutate(PercDisab = SumDisab/pop,
         PCT = ntile(PercDisab, 100)/100)

ValueP <- E_DISABL %>% 
  ggplot(aes(PercDisab))+
  geom_histogram(bins=10, fill='#69b3a2', color='white') + 
  scale_x_continuous(labels = scales::percent_format())+
  labs(#subtitle = "Percent of Population Estimated Below Poverty",
       y=NULL, x=NULL)

ValueP +
  plot_annotation(title = "Distribution Block Group Disabled Population")


```


## Single Parent Household

```{r, fig.width=12, fig.height=5, out.width="100%"}
# rm(longACS)
load(here::here("data", "E_SNGPNT.rda"))
objectNM <- unique(longACS$SVI_metric)
assign(objectNM, longACS, pos=1)

E_SNGPNT <-
  E_SNGPNT %>% 
  select(-vars) %>% 
  group_by_at(vars(-values)) %>%
  summarize(SumSngl = sum(values)) %>% 
  ungroup() %>% 
  left_join(., 
            E_HH %>% select(NAME, HH = values), 
            by = c("NAME")) %>% 
  mutate(PercHH_Sing= SumSngl/HH,
         PCT = ntile(PercHH_Sing, 100)/100)

ValueP <- E_SNGPNT %>% 
  ggplot(aes(PercHH_Sing))+
  geom_histogram(bins=10, fill='#69b3a2', color='white') + 
  scale_x_continuous(labels = scales::percent_format())+
  labs(#subtitle = "Percent of Population Estimated Below Poverty",
       y=NULL, x=NULL)

ValueP +
  plot_annotation(title = "Distribution Block Group Single Parents")


```

# Group: Minorty status {.tabset .tabset-pills}

## Minority

```{r, fig.width=12, fig.height=5, out.width="100%"}
# rm(longACS)
load(here::here("data", "E_MINRTY.rda"))
objectNM <- unique(longACS$SVI_metric)
assign(objectNM, longACS, pos=1)

E_MINRTY <-
  E_MINRTY %>% 
  left_join(., 
            E_TOTPOP %>% group_by(tract) %>% summarise(pop = sum(values)), 
            by = c("tract")) %>% 
  mutate(PercPop_Mino=  values/pop,
         PCT = ntile(PercPop_Mino, 100)/100)

ValueP <- E_MINRTY %>% 
  ggplot(aes(PercPop_Mino))+
  geom_histogram(bins=10, fill='#69b3a2', color='white') + 
  scale_x_continuous(labels = scales::percent_format())+
  labs(#subtitle = "Percent of Population Estimated Below Poverty",
       y=NULL, x=NULL)

ValueP +
  plot_annotation(title = "Distribution Tract Minority")


```

## Speak English "less than well"

```{r, fig.width=12, fig.height=5, out.width="100%"}
# rm(longACS)
load(here::here("data", "E_LIMENG.rda"))
objectNM <- unique(longACS$SVI_metric)
assign(objectNM, longACS, pos=1)

E_LIMENG <-
  E_LIMENG %>% 
  select(-vars) %>% 
  group_by_at(vars(-values)) %>%
  summarize(SumSpeak = sum(values)) %>% 
  ungroup() %>% 
  left_join(., 
            E_TOTPOP %>% select(NAME, pop = values), 
            by = c("NAME")) %>% 
  mutate(PercPop_Speak= SumSpeak/pop,
         PCT = ntile(PercPop_Speak, 100)/100)

ValueP <- E_LIMENG %>% 
  ggplot(aes(PercPop_Speak))+
  geom_histogram(bins=10, fill='#69b3a2', color='white') + 
  scale_x_continuous(labels = scales::percent_format())+
  labs(#subtitle = "Percent of Population Estimated Below Poverty",
       y=NULL, x=NULL)

ValueP +
  plot_annotation(title = "Distribution Block Group Speak English Less than Well")


```



# Group: Housing Type and Transportation {.tabset .tabset-pills}

## Multi-unit structures

```{r, fig.width=12, fig.height=5, out.width="100%"}
# rm(longACS)
load(here::here("data", "E_MUNIT.rda"))
objectNM <- unique(longACS$SVI_metric)
assign(objectNM, longACS, pos=1)

E_MUNIT <-
  E_MUNIT %>% 
  select(-vars) %>% 
  group_by_at(vars(-values)) %>%
  summarize(SumMultUN = sum(values)) %>% 
  ungroup() %>% 
  left_join(., 
            E_HU %>% select(NAME, HU = values), 
            by = c("NAME")) %>% 
  mutate(PercHU_MULT= SumMultUN/HU,
         PCT = ntile(PercHU_MULT, 100)/100)

ValueP <- E_MUNIT %>% 
  ggplot(aes(PercHU_MULT))+
  geom_histogram(bins=10, fill='#69b3a2', color='white') + 
  scale_x_continuous(labels = scales::percent_format())+
  labs(#subtitle = "Percent of Population Estimated Below Poverty",
       y=NULL, x=NULL)

ValueP +
  plot_annotation(title = "Distribution Block Group Multi Units")


```


## Mobile Homes

```{r, fig.width=12, fig.height=5, out.width="100%"}
# rm(longACS)
load(here::here("data", "E_MOBILE.rda"))
objectNM <- unique(longACS$SVI_metric)
assign(objectNM, longACS, pos=1)

E_MOBILE <-
  E_MOBILE %>% 
  # select(-vars) %>% 
  # group_by_at(vars(-values)) %>%
  # summarize(SumMultUN = sum(values)) %>% 
  # ungroup() %>% 
  left_join(., 
            E_HU %>% select(NAME, HU = values), 
            by = c("NAME")) %>% 
  mutate(PercHU_Mobl= values/HU,
         PCT = ntile(PercHU_Mobl, 100)/100)

ValueP <- E_MOBILE %>% 
  ggplot(aes(PercHU_Mobl))+
  geom_histogram(bins=10, fill='#69b3a2', color='white') + 
  scale_x_continuous(labels = scales::percent_format())+
  labs(#subtitle = "Percent of Population Estimated Below Poverty",
       y=NULL, x=NULL)

ValueP +
  plot_annotation(title = "Distribution Block Group Mobile Homes")


```

## Crowding

```{r, fig.width=12, fig.height=5, out.width="100%"}
# rm(longACS)
load(here::here("data", "E_CROWD.rda"))
objectNM <- unique(longACS$SVI_metric)
assign(objectNM, longACS, pos=1)

E_CROWD <-
  E_CROWD %>% 
  select(-vars) %>%
  group_by_at(vars(-values)) %>%
  summarize(SumCrowd= sum(values)) %>%
  ungroup() %>%
  left_join(., 
            E_HU %>% select(NAME, HU = values), 
            by = c("NAME")) %>% 
  mutate(PercHU_Crowd= SumCrowd/HU,
         PCT = ntile(PercHU_Crowd, 100)/100)

ValueP <- E_CROWD %>% 
  ggplot(aes(PercHU_Crowd))+
  geom_histogram(bins=10, fill='#69b3a2', color='white') + 
  scale_x_continuous(labels = scales::percent_format())+
  labs(#subtitle = "Percent of Population Estimated Below Poverty",
       y=NULL, x=NULL)

ValueP +
  plot_annotation(title = "Distribution Block Group Crowded Housing")


```

## No vehicles

```{r, fig.width=12, fig.height=5, out.width="100%"}
# rm(longACS)
load(here::here("data", "E_NOVEH.rda"))
objectNM <- unique(longACS$SVI_metric)
assign(objectNM, longACS, pos=1)

E_NOVEH <-
  E_NOVEH %>% 
  select(-vars) %>%
  group_by_at(vars(-values)) %>%
  summarize(SumNoVa= sum(values)) %>%
  ungroup() %>%
  left_join(., 
            E_HU %>% select(NAME, HU = values), 
            by = c("NAME")) %>% 
  mutate(PercHU_nova= SumNoVa/HU,
         PCT = ntile(PercHU_nova, 100)/100)

ValueP <- E_NOVEH %>% 
  ggplot(aes(PercHU_nova))+
  geom_histogram(bins=10, fill='#69b3a2', color='white') + 
  scale_x_continuous(labels = scales::percent_format())+
  labs(#subtitle = "Percent of Population Estimated Below Poverty",
       y=NULL, x=NULL)

ValueP +
  plot_annotation(title = "Distribution Block Group Housing with no Vehic")


```


## Group Quarters

```{r, fig.width=12, fig.height=5, out.width="100%"}
# rm(longACS)
load(here::here("data", "E_GROUPQ.rda"))
objectNM <- unique(longACS$SVI_metric)
assign(objectNM, longACS, pos=1)

E_GROUPQ <-
  E_GROUPQ %>% 
  left_join(., 
            E_TOTPOP %>% group_by(tract) %>% summarise(pop = sum(values)), 
            by = c("tract")) %>% 
  mutate(PercPop_GQ= values/pop,
         PCT = ntile(PercPop_GQ, 100)/100)

ValueP <- E_GROUPQ %>% 
  ggplot(aes(PercPop_GQ))+
  geom_histogram(bins=10, fill='#69b3a2', color='white') + 
  scale_x_continuous(labels = scales::percent_format())+
  labs(#subtitle = "Percent of Population Estimated Below Poverty",
       y=NULL, x=NULL)

ValueP +
  plot_annotation(title = "Distribution Block Group Housing with Group Quarters")


```



# Combine datasets

```{r}


CombinedSVI <-
  left_join(E_TOTPOP %>% select(-vars, -SVI_metric, -SVI_group, POP = values),
            E_HU %>% select(-vars, -SVI_metric, -SVI_group, HU = values),
              by = c("state", "county", "tract", "block_group", "NAME")) %>% 
  left_join(., 
            E_HH %>% select(-vars, -SVI_metric, -SVI_group, HH = values),
              by = c("state", "county", "tract", "block_group", "NAME")) %>% 
  # Pulling in the Socioexonomic Status related variables
  left_join(.,
            E_POV %>% select(-SVI_metric, -SVI_group, -pop, Count_Poverty = SumPov, Pctl_POV  = PCT),
              by = c("state", "county", "tract", "block_group", "NAME")) %>% 
  left_join(.,
            E_UNEMP %>% select(-SVI_metric, -SVI_group, -pop, -vars, Count_UnEMP = values, Pctl_unEMP  = PCT),
              by = c("state", "county", "tract", "block_group", "NAME"))  %>% 
  left_join(.,
            E_PCI %>% select(-SVI_metric, -SVI_group, -vars, PCI = values, Pctl_PCI  = PCT),
              by = c("state", "county", "tract", "block_group", "NAME"))   %>% 
  left_join(.,
            E_NOHSDP %>% select(-SVI_metric, -SVI_group, -NAME, -pop, Count_NoHSDipl = SumNoDP, Pctl_NoHSDipl  = PCT),
              by = c("state", "county", "tract")) %>%
  # Pulling in the Household Composition and Disability related variables  
  left_join(.,
            E_AGE65 %>% select(-SVI_metric, -SVI_group, -pop,  Count_Older65 = SumOlder, Pctl_Older65  = PCT),
              by = c("state", "county", "tract", "block_group", "NAME"))  %>% 
  left_join(.,
            E_AGE17 %>% select(-SVI_metric, -SVI_group, -pop, Count_Young17 = SumYoung, Pctl_Young17  = PCT),
              by = c("state", "county", "tract", "block_group", "NAME")) %>% 
  left_join(.,
            E_DISABL %>% select(-SVI_metric, -SVI_group, -NAME, -pop, Count_Disabled = SumDisab, Pctl_Disabled  = PCT),
              by = c("state", "county", "tract")) %>%
  left_join(.,
            E_SNGPNT %>% select(-SVI_metric, -SVI_group, -HH, Count_SnglPrnt = SumSngl, Pctl_SnglPrnt  = PCT),
              by = c("state", "county", "tract", "block_group", "NAME")) %>% 
  # Pulling in  Minority Status and Language
  left_join(.,
            E_MINRTY %>% select(-SVI_metric, -SVI_group, -NAME,-vars,  -pop, Count_Minority = values, Pctl_Minority  = PCT),
              by = c("state", "county", "tract")) %>%
  left_join(.,
            E_LIMENG %>% select(-SVI_metric, -SVI_group, -pop, Count_LessEnglish = SumSpeak, Pctl_LessEnglish  = PCT),
              by = c("state", "county", "tract", "block_group", "NAME")) %>% 
  # Pulling in Housing Type and Transportation
  left_join(.,
            E_MUNIT %>% select(-SVI_metric, -SVI_group, -HU, Count_MultUnit = SumMultUN, Pctl_MultUnit  = PCT),
              by = c("state", "county", "tract", "block_group", "NAME")) %>% 
  left_join(.,
            E_MOBILE %>% select(-SVI_metric, -SVI_group, -vars, -HU, Count_MobileH = values, Pctl_MobileH  = PCT),
              by = c("state", "county", "tract", "block_group", "NAME")) %>% 
  left_join(.,
            E_CROWD %>% select(-SVI_metric, -SVI_group, -HU, Count_Crowd = SumCrowd, Pctl_Crowd  = PCT),
              by = c("state", "county", "tract", "block_group", "NAME")) %>% 
  left_join(.,
            E_GROUPQ %>% select(-SVI_metric, -SVI_group, -NAME, -pop, -vars, Count_GroupQtr = values, Pctl_GroupQtr  = PCT),
              by = c("state", "county", "tract"))



save(CombinedSVI, file = here::here("data", "CombinedSVI.rda"))


```












